---
title: "Buoy_Hyungu Lee"
format: pdf
editor: visual
---

## Packages to Install

```{r, warning = FALSE, message=FALSE}
library(data.table)
library(dplyr)
library(lubridate)
library(ggplot2)
library(zoo)
library(tibble)
library(readr)
```

Getting Buoy Data

```{r, warning=FALSE, message=FALSE}
file_root <- "https://www.ndbc.noaa.gov/view_text_file.php?filename=44013h"
tail <- ".txt.gz&dir=data/historical/stdmet/"

load_buoy_data1 <- function(year) {
  path <- paste0(file_root, year, tail)
  
  
  if (year < 2007) {
  header <- scan(path, what = 'character', nlines = 1)
  buoy <- read.table(path, fill = TRUE, header = TRUE, sep = "")
  buoy <- add_column(buoy, mm = NA, .after = "hh")
  buoy <- add_column(buoy, TIDE = NA, .after = "VIS")
    
  } else {
  header <- scan(path, what = 'character', nlines = 1)  
  buoy <- fread(path, header = FALSE, skip = 1, fill = TRUE)

    setnames(buoy, header)
  }
  
  #return(buoy)
}

all_data1 <- lapply(1985:2024, load_buoy_data1)

combined_data1 <- rbindlist(all_data1, fill = TRUE)
```

```{r}
load_buoy_data <- function(year) {
  path <- paste0(file_root, year, tail)

  header <- scan(path, what = 'character', nlines = 1)
  num_columns <- length(header)  

  if (num_columns == 16) {
    buoy <- read.table(path, fill = TRUE, header = TRUE, sep = "")
    buoy <- add_column(buoy, mm = NA, .after = "hh")
    buoy <- add_column(buoy, TIDE = NA, .after = "VIS")

  } else if (num_columns == 17) {
    buoy <- read.table(path, fill = TRUE, header = TRUE, sep = "")
    buoy <- add_column(buoy, TIDE = NA, .after = "VIS")

  } else {
    buoy <- fread(path, header = FALSE, skip = 1, fill = TRUE)
    setnames(buoy, header)
  }

  return(buoy)
}
all_data <- lapply(1985:2024, load_buoy_data)
combined_data <- rbindlist(all_data, fill = TRUE)
```

```{r}
combined_data1 <- combined_data1 %>%
  mutate(
    YY = as.character(YY),
    `#YY` = as.character(`#YY`),
    YYYY = as.character(YYYY)
  )

# Combine year columns safely using coalesce
combined_data1 <- combined_data1 %>%
  mutate(YYYY = coalesce(YYYY, `#YY`, YY))
combined_data1 <- combined_data1 %>%
  mutate(BAR = coalesce(as.numeric(BAR), as.numeric(PRES)),  # Convert BAR and PRES to numeric
    WD = coalesce(as.numeric(WD), as.numeric(WDIR)))

combined_data1 <- combined_data1 %>%
  select(-TIDE, -TIDE.1, -mm,- WDIR, -PRES,-`#YY`,-YY)

combined_data1$datetime <- ymd_h(paste(combined_data1$YYYY, combined_data1$MM, combined_data1$DD, combined_data1$hh, sep = "-"))

combined_data1 <- combined_data1 %>%
  mutate(across(everything(), 
                ~ na_if(as.numeric(as.character(.)), 99) %>%
                na_if(999) %>%
                na_if(9999)))

#summary(combined_data)
#str(combined_data)
#str(combined_data$datetime)
if (!inherits(combined_data1$datetime, "POSIXct")) {
  combined_data1$datetime <- ymd_h(paste(combined_data1$YYYY, combined_data1$MM, combined_data1$DD, combined_data1$hh, sep = "-"))
}
```

## Compiling all years' data

1.  Try to remember why 2007 is the split year
2.  Why are using different functions to read data prior and post 2007?
3.  What are some new function you got to know from following code? what do these functions do?

```{r, warning=FALSE, message=FALSE}
file_root <- "https://www.ndbc.noaa.gov/view_text_file.php?filename=44013h"
tail <- ".txt.gz&dir=data/historical/stdmet/"

load_buoy_data <- function(year) {
  path <- paste0(file_root, year, tail)

  header <- scan(path, what = 'character', nlines = 1)
  num_columns <- length(header)  

  if (num_columns == 16) {
    buoy <- read.table(path, fill = TRUE, header = TRUE, sep = "")
    buoy <- add_column(buoy, mm = NA, .after = "hh")
    buoy <- add_column(buoy, TIDE = NA, .after = "VIS")

  } else if (num_columns == 17) {
    buoy <- read.table(path, fill = TRUE, header = TRUE, sep = "")
    buoy <- add_column(buoy, TIDE = NA, .after = "VIS")

  } else {
    buoy <- fread(path, header = FALSE, skip = 1, fill = TRUE)
    setnames(buoy, header)
  }

  return(buoy)
}
all_data <- lapply(1985:2024, load_buoy_data)
combined_data <- rbindlist(all_data, fill = TRUE)
```

## Cleaning and Organizing the data

We start by merging all different version of `Year` column. We do same with other columns which are same but having data for certain set of years. We remove the remaining columns after merging them.

Creating datetime column using `lubridate()`

```{r}
combined_data <- combined_data %>%
  mutate(
    YY = as.character(YY),
    `#YY` = as.character(`#YY`),
    YYYY = as.character(YYYY)
  )

# Combine year columns safely using coalesce
combined_data <- combined_data %>%
  mutate(YYYY = coalesce(YYYY, `#YY`, YY))
combined_data <- combined_data %>%
  mutate(BAR = coalesce(as.numeric(BAR), as.numeric(PRES)),  # Convert BAR and PRES to numeric
    WD = coalesce(as.numeric(WD), as.numeric(WDIR)))

combined_data <- combined_data %>%
  select(-TIDE, -TIDE.1, -mm,- WDIR, -PRES,-`#YY`,-YY)

combined_data$datetime <- ymd_h(paste(combined_data$YYYY, combined_data$MM, combined_data$DD, combined_data$hh, sep = "-"))

combined_data <- combined_data %>%
  mutate(across(everything(), 
                ~ na_if(as.numeric(as.character(.)), 99) %>%
                na_if(999) %>%
                na_if(9999)))

#summary(combined_data)
#str(combined_data)
#str(combined_data$datetime)
if (!inherits(combined_data$datetime, "POSIXct")) {
  combined_data$datetime <- ymd_h(paste(combined_data$YYYY, combined_data$MM, combined_data$DD, combined_data$hh, sep = "-"))
}
```

```{r}
combined_data <- combined_data %>%
  mutate(Year = year(datetime),
         Month = month(datetime),
         Day = day(datetime),
         Hour = hour(datetime))

#How has the annual average wind speed at buoy 44013 fluctuated from #1985 to 2024, and does the data indicate a consistent long-term trend #or is it characterized more by year-to-year variability?

# Plot 1: Trend of Annual Average Wind Speed
plot1_avg_wind_speed <- ggplot(yearly_wind_analysis, aes(x = Year, y = avg_wspd)) +
  geom_line(color = "navy", size = 1) +
  geom_point(color = "navy") +
  labs(
    title = "Annual Average Wind Speed (WSPD) Trend",
    x = "Year",
    y = "Average Wind Speed (m/s)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot1_avg_wind_speed)
```

## From the graph, we can make the following observations:

-   Significant Year-to-Year Variability: The average wind speed fluctuates considerably from one year to the next. Some years were significantly windier than others.

-   Absence of a Clear Long-Term Trend: There is no obvious, consistent long-term trend of either increasing or decreasing wind speed over the entire period. Instead, the graph shows cyclical patterns where wind speeds rise and fall over several years.

-   Peak Wind Speeds: The highest average wind speed was recorded around 1990, when it exceeded 7.0 m/s. Other relatively windy periods occurred in the late 1990s, around 2010, and around 2020.

-   Lowest Wind Speeds: The lowest point on the graph appears in the period after 2020, where the average wind speed dropped below 5.5 m/s. Another period of low wind speed occurred just before the year 2000.
